\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2015-12-28 overview.tex
minor edits

The goal of our work is to find distributed system Heisenbugs \emph{before} they reach production. Typical distributed systems consist of multiple components that interact with each other via \emph{message passing}. If messages---or unexpected failures and timeouts---are not handled properly, they can lead to subtle bugs. To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{specifying} properties of \emph{correctness}, and \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, client requests, failures and timeouts.

Modeling using \psharp involves three core activities. First, the developer must modify the original system so that messages are not sent through the real network, but are instead dispatched through the \texttt{PSharp.Send(...)} method. Such modification does not need to be invasive, as it can be performed using virtual method dispatch, a \csharp language feature widely used for testing. Second, the developer must write a \psharp \emph{test harness} that drives the system towards interesting behaviors, and nondeterministically triggers various distributed events (see \S\ref{sec:overview:model}). The harness is essentially an environment model of the system. The purpose of these first two activities is to explicitly declare all sources of nondeterminism in the system using \psharp. Finally, the developer must specify the criteria for correctness of an execution of the system-under-test. Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}).

During testing, the \psharp runtime is aware of all sources of nondeterminism that were declared during modeling, and exploits this knowledge to create a scheduling point each time a nondeterministic choice has to be taken. The \psharp testing engine will serialize (as single-box) the system, and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a user-supplied bound (e.g. in number of executions or time), or it hits a safety or liveness property violation. This testing process is fully automatic, has no false-positives (assuming an accurate test harness), and can reproduce found bugs by replaying buggy schedules. After a bug is discovered, \psharp generates a trace that represents the buggy schedule. In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events, and thus is easier to debug.

@@

In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled by \psharp environment. The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space. Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness and environment using \psharp. However, we argue that developers already spend significant time in building test suites and testing a distributed system prior to deployment. The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}).

@@

To systematically test the example of Figure~\ref{fig:example}, the developer must first create a \psharp test harness, and then specify the correctness properties of the system. Figure~\ref{fig:example:model} illustrates a test harness that can find the two bugs discussed in \S\ref{sec:overview:example}. Each box in the figure represents a concurrently running \psharp machine, while an arrow represents an event being send from one machine to another.

We did not model the server and SN components, so we can test their actual implementation. Each real component is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component. We decided to model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}. We also modeled the timer so that \psharp takes control of all time-related nondeterminism in the system. This allows the \psharp runtime to control when a \texttt{Timeout} event will be sent to the SNs during testing, and systematically explore different schedules.

@@

The first bug in the example of \S\ref{sec:overview:example} is a safety bug. To find it, the developer can write a safety monitor (see Figure~\ref{fig:example:model}) that contains a map from unique SN ids to a boolean value, which denotes if the SN is a replica or not. Each time an SN replicates the latest data, it notifies the monitor to update the map. Each time the server issues an \texttt{Ack}, it also notifies the monitor. If the monitor detects that an \texttt{Ack} was sent without three replicas actually existing, a safety violation is triggered.

>>>>

The goal of our work is to find bugs in distributed system \emph{before} they reach production. Typical distributed systems consist of multiple components that interact with each other via \emph{message passing}. If messages---or unexpected failures and timeouts---are not handled properly, they can lead to subtle bugs. To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{specifying} properties of \emph{correctness}, and \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, client requests, failures and timeouts.

Modeling using \psharp involves three core activities. First, the developer must modify the original system so that messages are not sent through the real network, but are instead dispatched through the \texttt{PSharp.Send(...)} method. Such modification does not need to be invasive, as it can be performed using virtual method dispatch, a \csharp language feature widely used for testing. Second, the developer must write a \psharp \emph{test harness} that drives the system towards interesting behaviors, and nondeterministically triggers various events (see \S\ref{sec:overview:model}). The harness is essentially a model of the environment of the system. The purpose of these first two activities is to explicitly declare all sources of nondeterminism in the system using \psharp. Finally, the developer must specify the criteria for correctness of an execution of the system-under-test. Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}).

During testing, the \psharp runtime is aware of all sources of nondeterminism that were declared during modeling, and exploits this knowledge to create a scheduling point each time a nondeterministic choice has to be taken. The \psharp testing engine will serialize (in a single-box) the system, and repeatedly execute it from start to completion, each time exploring a potentially different set of nondeterministic choices, until it either reaches a user-supplied bound (e.g. in number of executions or time), or it hits a safety or liveness property violation. This testing process is fully automatic, has no false-positives (assuming an accurate test harness), and can reproduce found bugs by replaying buggy schedules. After a bug is discovered, \psharp generates a trace that represents the buggy schedule. In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events, and thus is easier to debug.

@@

In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled \psharp environment. The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space. Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp. However, we argue that developers already spend significant time in building test suites for distributed systems prior to deployment. The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}).

@@

To systematically test the example of Figure~\ref{fig:example}, the developer must first create a \psharp test harness, and then specify the correctness properties of the system. Figure~\ref{fig:example:model} illustrates a test harness that can find the two bugs discussed in \S\ref{sec:overview:example}. Each box in the figure represents a concurrently running \psharp machine, while an arrow represents an event being sent from one machine to another.

We do not model the server and SN components since we want to test their actual implementation. Each real component is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component. We model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}. We also model the timer so that \psharp takes control of all time-related nondeterminism in the system. This allows the \psharp runtime to control when a \texttt{Timeout} event will be sent to the SNs during testing, and systematically explore different schedules.

@@

The first bug in the example of \S\ref{sec:overview:example} is a safety bug. To find it, the developer can write a safety monitor (see Figure~\ref{fig:example:model}) that contains a map from unique SN ids to a Boolean value, which denotes if the SN is a replica or not. Each time an SN replicates the latest data, it notifies the monitor to update the map. Each time the server issues an \texttt{Ack}, it also notifies the monitor. If the monitor detects that an \texttt{Ack} was sent without three replicas actually existing, a safety violation is triggered.
